---
title: "Delta AUC Nested Model – Complete Analysis"
author: "Yuxuan Liu"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: cosmo
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 10, fig.height = 6)
```

# Setup

## Load Packages

```{r packages}
# Determine project root (works in both RStudio and command-line knit)
script_dir <- tryCatch(
  dirname(rstudioapi::getActiveDocumentContext()$path),
  error = function(e) NULL
)
if (is.null(script_dir) || script_dir == "") {
  # When knitting from command line, knitr sets the working directory to
  # the directory containing the Rmd file.
  script_dir <- getwd()
}
cat("Project directory:", script_dir, "\n")

source(file.path(script_dir, "R", "00_packages.R"))
```

## Load Function Modules

```{r source-modules}
source(file.path(script_dir, "R", "01_model_functions.R"))
source(file.path(script_dir, "R", "02_ground_truth.R"))
source(file.path(script_dir, "R", "03_simulation_RxB.R"))
source(file.path(script_dir, "R", "04_inference_viz.R"))
source(file.path(script_dir, "R", "05_loo_cv.R"))
source(file.path(script_dir, "R", "06_real_data_ptb.R"))
source(file.path(script_dir, "R", "07_real_data_heart.R"))
```

## Simulation Parameters

> **Note:** All simulation sizes (R, B, n) are kept deliberately small so
> this document renders in a few minutes. For production-quality results,
> increase `R` to 10 000 for ground truth, `R=50, B=20` for R×B simulation,
> and restore the full sample sizes (n=2000) – see `demo_run_all.R` for
> the original values.

```{r params}
p   <- 100   # number of noise features
mu  <- 1     # signal mean shift
sig <- 10    # noise SD
snr <- (mu - 0)^2 / sig^2
cat(sprintf("Signal-to-Noise Ratio (SNR) = %.4f\n", snr))
```

---

# Ground Truth

We estimate the true $\Delta$AUC (full − reduced) via large-scale Monte Carlo
with independent training and test sets.

## Setting 1: $n_{train} = 500$, $n_{test} = 1000$

```{r ground-truth-1, cache=TRUE}
# NOTE: R kept small for demo speed. Increase R for production runs.
ground_truth_1 <- compute_ground_truth(
  n1 = 250, n2 = 250,
  n1_test = 500, n2_test = 500,
  R = 100,
  p = p, mu = mu, sig = sig,
  model_types = c("lda", "logistic")
)
```

```{r gt1-print}
cat("Ground truth setting 1 (n_train=500, n_test=1000):\n")
print(ground_truth_1$ground_truth)
```

## Setting 2: $n_{train} = 400$, $n_{test} = 1000$

```{r ground-truth-2, cache=TRUE}
ground_truth_2 <- compute_ground_truth(
  n1 = 200, n2 = 200,
  n1_test = 500, n2_test = 500,
  R = 100,
  p = p, mu = mu, sig = sig,
  model_types = c("lda", "logistic")
)
```

```{r gt2-print}
cat("Ground truth setting 2 (n_train=400, n_test=1000):\n")
print(ground_truth_2$ground_truth)
```

---

# R × B Simulation

Non-split model is fit **once** per Monte Carlo replicate (deterministic on
same data). Only the sample-split evaluation is repeated B times.
A single iteration is timed first so you can tune R and B.

```{r sim-rxb, cache=TRUE}
# NOTE: R, B kept small for demo speed. Increase for production.
sim_result <- run_simulation_RxB(
  n1 = 250, n2 = 250,
  R  = 10, B = 10,
  p  = p, mu = mu, sig = sig,
  test_frac   = 0.25,
  model_types = c("lda", "logistic"),
  nrounds     = 100,
  reduced_formula = "Y ~ X1"
)
```

```{r sim-timing}
cat(sprintf("Single iteration time : %.2f sec\n", sim_result$time_single))
cat(sprintf("Total simulation time : %.1f sec (%.1f min)\n",
            sim_result$time_total, sim_result$time_total / 60))
```

---

# Inference & Visualization

Inference is computed over all **R × B** split samples (pooled).
Coverage rates are computed against ground truth setting 1.

```{r inference}
inference <- compute_inference(sim_result,
                               ground_truth = ground_truth_1$ground_truth)
print_inference_summary(inference)
```

## AUC Difference Histograms

```{r hist-auc, fig.height=10}
plot_auc_histograms(sim_result, xlim_range = c(-0.15, 0.15))
```

---

# Leave-One-Out Cross Validation

LOO-CV simulation for LDA and Logistic regression.
Each Monte Carlo replicate generates fresh data; non-split and LOO results
are both deterministic for a given dataset.

```{r loo-cv, cache=TRUE}
# NOTE: R kept small for demo. LOO-CV is expensive (refits per observation).
loo_result <- run_loo_cv_simulation(
  n1 = 100, n2 = 100,
  R  = 3,
  p  = p, mu = mu, sig = sig,
  reduced_formula = "Y ~ X1"
)
```

```{r loo-summary}
print_loo_summary(loo_result)
```

## LOO-CV Histograms

```{r loo-hist}
plot_loo_histograms(loo_result)
```

---

# Real Data Application – Preterm Birth

Uses the `ptb` dataset from the **SELR** package.

- **Scenario 1**: Full model (all features + interactions) vs reduced model.
- **Scenario 2**: HC-only features vs base model.

Non-split results are computed once; split results repeated R times.

```{r ptb-analysis, cache=TRUE}
if (exists("ptb")) {
  ptb_result <- run_ptb_analysis(
    R = 50,
    model_types = c("lda", "logistic"),
    nrounds = 100
  )
} else {
  cat("PTB dataset not found. Skipping.\n")
  cat("Make sure the SELR package is installed and loaded.\n")
  ptb_result <- NULL
}
```

```{r ptb-summary}
if (!is.null(ptb_result)) {
  print_ptb_summary(ptb_result)
}
```

---

# Real Data Application – Framingham Heart Study

Uses the `framingham` dataset from the **riskCommunicator** package.

- Full model: `Y ~ BMI + SYSBP + DIABP + TOTCHOL + AGE`
- Reduced model: `Y ~ BMI + SYSBP + DIABP + TOTCHOL` (drops AGE)

```{r heart-analysis, cache=TRUE}
if (exists("framingham")) {
  heart_result <- run_heart_analysis(R = 50)
} else {
  cat("Framingham dataset not found. Skipping.\n")
  cat("Make sure the riskCommunicator package is installed and loaded.\n")
  heart_result <- NULL
}
```

```{r heart-summary, fig.height=5}
if (!is.null(heart_result)) {
  print_heart_summary(heart_result)
}
```

---

# Session Info

```{r session-info}
sessionInfo()
```
